---
title: "courts"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rgdal)
library(dfoptim)
source("helpers.R")
```

# Assumptions
## General Assumptions
- Assumptions due to limited data:
 
  - Satellite courts lack data because the courthouse hours MAG dataset is not detailed to the level of the court asset MAG dataset, therefore
    - Satellite courts are assumed to service the same set of business lines as the base court in the CD
    - Satellite courts are assumed to have 1 courtroom
  
  - Assume individuals appear at courthouses within their CD of home origin due to limited information on where court demand is coming from
 
  - Assume that a portion of of individuals are "assigned" and another portion chooses closeset available courtroom due to limited information on the number of trips per individual and where court demand is coming from
    - This split is defaulted to 20/80 for civil, family, small claims and 10/90 for OCJcrim, SCJcrim
  
  - Four years of back data may not be enough to tease out future trends where hrs/case is increasing and may cause an inflection in forecast years - future extended data set may allow for various demand assessment approaches and/or more accurate rate production from historical data
  
## Processing Assumptions

  - CD 3553 (Greater Sudbury) is within CD 3552 (Sudbury) with the courts located in Greater Sudbury. Therefore, the two CDs are combined and treated as one CD due to limited on information on where individuals attending court are coming from

## Import Data

```{r import data}
# Improt TRESO information to define system
treso_shp <- readOGR(dsn ="input/treso", layer = "TRESO_Zones_SocioData_Gatineau_LCC")
treso_zone_system <- read.csv("input/treso/treso_zone_system.csv")


# Import TRESO Travel TIme Skims to distribute demand to courts
travel_time_skim <- readRDS("output/treso/treso_travel_time.rds")


# Import prepared TRESO projection population with age groups segmented to enable MAG business line (casetype) breakdown
treso_population <- readRDS("input/mag/treso_population_mag.rds")

# Import courthouse hours dataset recieved from MAG
court_hours <- readRDS("cache/mag/courthouse_hours.rds")

# Import courthouse asset lists that includes satellite & base courts from MAG
courthouse_asset <- readRDS("cache/mag/courthouse_master.rds")
```

## Data Preparation

```{r data preparation}
# Gather the TRESO Population table into long format ro prepare data for business line, year, and TRESO Zone goruping
treso_population_spread <- treso_population %>% 
  select(-csduid) %>% 
  rename_at(vars(contains("population")), list(~substr(., 12, 15))) %>% 
  gather(key="year", value = "population", -treso_zone, -age.group)

# Transform age ranges to business line populations by zone and projection year to sum zone population by matched business line age ranges
treso_population_bl <- treso_population_spread %>% 
  mutate(OCJcrim = ifelse(age.group %in% c("12-17", "18", "19", "20-49"), population, 0)) %>% 
  mutate(SCJcrim = ifelse(age.group %in% c("18", "19", "20-49"), population, 0)) %>% 
  mutate(OCJfam = ifelse(age.group %in% c("0-11", "12-17"), population, 0)) %>% 
  mutate(SCJfam = ifelse(age.group %in% c("0-11", "12-17", "18", "19", "20-49", "50-59"), population, 0)) %>% 
  mutate(civil = ifelse(age.group %in% c("20-49", "50-59", "60-69"), population, 0)) %>%
  mutate(smac = ifelse(age.group %in% c("20-49", "50-59", "60-69"), population, 0)) %>%
  group_by(treso_zone, year) %>% 
  summarise_if(is.numeric, sum) %>% 
  left_join(select(treso_zone_system, treso_id, cduid, cdname), by = c("treso_zone" = "treso_id"))

# Output the population by business line as long format for historical court hour matching
treso_population_bl_long <- treso_population_bl %>%
  gather(casetypes, population, OCJcrim:smac) %>%
  arrange(treso_zone, casetypes , year) %>% 
  ungroup()

saveRDS(treso_population_bl_long, "cache/mag/treso_population_bl_long.rds")

```


```{r combine hours with population}

# Read the cached treso_population_bl_long table from the previous data preparation code chunk
treso_population_bl_long <- readRDS("cache/mag/treso_population_bl_long.rds")

# Summarize to CD level for population due to court hours being at the base court CD level, then
# Combine with hours 
treso_population_bl_cd <- treso_population_bl_long %>% 
  group_by(year, cduid, cdname, casetypes) %>%
  summarise(population = sum(population)) %>% 
  ungroup()

court_hours_population <- court_hours %>% 
  left_join(treso_population_bl_cd, by = c('year', 'casetypes', 'cduid', 'cdname')) %>% 
  mutate_at("year", as.numeric) %>% 
  rename(hours = value, treso_id = treso.id.pos) %>% 
  select(-lat, -long, -csduid, -csdname, - courthouse.lat, - courthouse.long)

# IMPORTANT
# CD 3553 (Greater Sudbury) is within CD 3552 (Sudbury)
# The courts are located in Greater Sudbury, so the two CDs are combined and treated as one CD
sudbury_population_bl_cd <- treso_population_bl_cd %>% 
  mutate_at("year", as.numeric) %>% 
  filter(cduid == 3552) %>% 
  mutate(cduid = 3553) %>% 
  rename(population.sudbury = population) %>% 
  select(year, cduid, casetypes, population.sudbury)
  
court_hours_population <- court_hours_population %>% 
  left_join(sudbury_population_bl_cd, by = c("cduid", "year", "casetypes")) %>% 
  mutate(population = ifelse(is.na(population.sudbury), population, population + population.sudbury))

cat(paste0("Quick check that the population in", "\n",
           "CD 3553 and 3552 is: ", 
           sum(filter(treso_population_bl_cd, year == 2018, cduid %in% c(3552, 3553))$population), "\n",
           "Combined is: ",
           (filter(court_hours_population, year == 2018, cduid == 3553) %>% group_by(bid, cduid) %>% summarise(population = sum(population)))$population[1]))

```

# Rate Calculation
## Base Year Rates

The age ranges for business lines are based on background information received as follows:

- OCJCrim: 12 - 49
- SCJCrim: 18 - 49
- OCJFam: 0 - 17
- SCJFam: 0 - 59
- Civil: 20 - 69
- Smac: 20 - 69

```{r population rates for base year 2018}
# Establish rate of business line courthours by population based on historical data. Available historical data is from 2014-2018.

BASE_YEAR <- 2018
WEIGHT_FACTOR <- "five_years" # Option between: (1)an average weighted approach ("five_years") (2)most recent historical data year (3)user custom weighting
WEIGHT_CUSTOM = tibble("year" = c(2018, 2017, 2016, 2015, 2014),
                       "weight" = c(1, 1, 1, 1, 1)) #For custom historical data weighting - default is equal weighting, equivelant to the five_years option


if (WEIGHT_FACTOR == "five_years") {
  # Equally weight the five past years to calculate population rates
  court_rate <- court_hours_population %>% 
    filter(year >= (BASE_YEAR - 5)) %>%
    group_by(bid, name, cduid, cdname, casetypes, year) %>% 
    # Population is duplicated for each bid, so don't sum it up
    summarise(population = first(population), hours = sum(hours)) %>% 
    mutate(pop.rate = hours / population) %>% 
  # Average the years, if there are differences
    group_by(bid, name, cduid, cdname, casetypes) %>% 
    summarise(pop.rate = mean(pop.rate),
              hours = sum(hours))
    
} else if (WEIGHT_FACTOR == "recent_year") {
  # Only use the previous year to calculate population rates
  court_rate <- court_hours_population %>% 
    filter(year >= (BASE_YEAR)) %>%
    group_by(bid, name, cduid, cdname, casetypes, year) %>% 
    summarise(population = sum(population), hours = sum(hours)) %>% 
    mutate(pop.rate = hours / population) %>% 
  # Average the years, if there are differences
    group_by(bid, name, cduid, cdname, casetypes) %>% 
    summarise(pop.rate = mean(pop.rate),
              hours = sum(hours))
  
} else if (WEIGHT_FACTOR == "custom") {
  # Use a custom weighting system
  court_rate <- court_hours_population %>% 
    left_join(WEIGHT_CUSTOM, by = "year") %>% 
    filter(!is.na(weight)) %>%
    group_by(bid, name, cduid, cdname, casetypes, year) %>% 
    summarise(population = sum(population), hours = sum(hours)) %>% 
    mutate(pop.rate = hours / population) %>% 
    # Average the years, if there are differences
    group_by(bid, name, cduid, cdname, casetypes) %>% 
    summarise(pop.rate = weighted.mean(pop.rate, weight),
              hours = sum(hours))
}

baseyear_rates <- court_rate %>% 
  # Add the contribution of the multiple courthouses together
  group_by(cduid, cdname, casetypes) %>% 
  summarise(pop.rate = sum(pop.rate)) %>% 
  replace_na(list(pop.rate = 0)) %>% 
  ungroup()
saveRDS(baseyear_rates, "cache/mag/baseyear_rates.rds")
```

## Projection Cases

```{r project cases by business lines}
# Apply established rates in previous chunk from cached baseyear_rates.rds table

future_cases_bl <- readRDS("cache/mag/baseyear_rates.rds") %>%
  left_join(treso_population_bl_cd, by = c("cduid", "cdname", "casetypes")) %>%
  spread(key = "year", value = "population") %>%
  select(-`2011`:-`2017`) %>%
  mutate_at(vars(starts_with("20")), list(~(. * pop.rate))) %>% 
  gather(year, court.hours, `2018`:`2041`)

saveRDS(future_cases_bl, "cache/mag/future_cases_bl.rds")

# Sanity check - this will miss population from Sudbury, so it will be slightly lower than observed
# But in the TRESO level which is what is actually used in modelling - Sudbury will share the rates of
# Greater Sudbury
print("If the 'recent_year' flag is used to calculate the baseyear_rate, then the court hours in 2018 should match")

print(filter(future_cases_bl, year == 2018) %>% group_by(cduid, cdname, casetypes) %>% summarise(court.hours = sum(court.hours)))

print(filter(court_hours_population, year == 2018) %>% group_by(cduid, cdname, casetypes) %>% summarise(court.hours = sum(hours)))

print(paste0("The calculated court hours is: ", (sum(filter(future_cases_bl, year == 2018)$court.hours)),
             " While the historical 2018 court hours is: ", (sum(filter(court_hours_population, year == 2018)$hours))))

# Calculate the cases at TRESO level
future_cases_bl_treso <- readRDS("cache/mag/treso_population_bl_long.rds")
baseyear_rates <- readRDS("cache/mag/baseyear_rates.rds")

# CD 3546 (Haliburton) is missing court hours, CD 3552 (Sudbury) will share rates from CD 3553 (Greater Sudbury)
# The case rates were calculated using the combined population
sudbury_rates <- baseyear_rates %>% 
  filter(cduid == 3553) %>% 
  select(cduid, casetypes, pop.rate) %>% 
  mutate(cduid = 3552) %>% 
  rename(pop.rate.sudbury = pop.rate)

future_cases_bl_treso <- future_cases_bl_treso %>% 
  left_join(baseyear_rates, by = c("cduid", "cdname", "casetypes")) %>% 
  left_join(sudbury_rates, by = c("cduid", "casetypes")) %>% 
  mutate(pop.rate = ifelse(!is.na(pop.rate.sudbury), pop.rate.sudbury, pop.rate)) %>% 
  select(-pop.rate.sudbury) %>% 
  mutate(court.hours = population * pop.rate) %>% 
  # TODO remove drop_na once hours for Haliburton is obtained
  drop_na(court.hours) 

# Sanity Check
print(paste0("The calculated court hours is: ", (sum(filter(future_cases_bl_treso, year == 2018)$court.hours)),
             " While the historical 2018 court hours is: ", (sum(filter(court_hours_population, year == 2018)$hours))))

saveRDS(future_cases_bl_treso, "cache/mag/future_cases_bl_treso.rds")



```

# Courtroom Required

```{r courtroom required}
# Calculate courtrooms required using business line specfic utilization characteristics from background data recieved

MODEL_YEAR <- 2018
BASE_YEAR <- 2018
OP_DAYS <- 249
# NEW_COURTROOM_SIZE <- 11500

# Build utilization table for each business line
utilization_rate <- tibble(casetypes = c('OCJcrim','SCJcrim','OCJfam','SCJfam','civil','smac'),
                           base.hours.dy = c(3.5, 2.6, 2.7, 2.7, 2.4, 2.7),
                           inflect.hours.dy = c(4.4, 3.7, 4.3, 4.1, 3.5, 4.0),
                           max.hours.dy = c(4.5, 3.8, 4.4, 4.3, 3.6, 4.1),
                           inflect.hours = c(15000, 9000, 5000, 6000, 5000, 6000),
                           max.hours = c(30000, 18000, 10000, 12000, 10000, 12000)
                           )

saveRDS(utilization_rate, "cache/mag/utilization_rate.rds")

# Apply user-defined 

# Calculate future court hours needed based on the business line specific utilization characteristics
future_needs <- readRDS("cache/mag/future_cases_bl.rds") %>%
  filter(year == MODEL_YEAR) %>% 
  left_join(utilization_rate, by = "casetypes") %>% 
  mutate(op.utilization = 0) %>% 
  mutate(op.utilization = ifelse(court.hours <= inflect.hours,
                              base.hours.dy + (inflect.hours.dy - base.hours.dy) / (inflect.hours - 0 ) * court.hours,
                              op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours > inflect.hours & court.hours <= max.hours,
                                 inflect.hours.dy + ((max.hours.dy - inflect.hours.dy) * (court.hours - inflect.hours)                                                       / (max.hours - inflect.hours)),
                                 op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours > max.hours,
                                 max.hours.dy,
                                 op.utilization)) %>% 
  mutate(courtrooms.needed = round(court.hours / (op.utilization * OP_DAYS), digits = 2))

saveRDS(future_needs, "cache/mag/future_needs.rds")
           
  
# Summarise to CD level
future_needs_cd <- future_needs %>% 
  group_by(cduid, cdname) %>% 
  summarise(courthours.needed = sum(court.hours),
            courtrooms.needed = sum(courtrooms.needed)) 

# Summarise courtrooms needed with the existing 2018 courtrooms dataset
courthouse_asset_with_projected_cd <- courthouse_asset %>%
  group_by(cduid, cdname) %>% 
  summarise(number.of.courtrooms.2018 = sum(courtrooms),
            rentable.square.feet.2018 = sum(rentable.square.feet)) %>% 
  left_join(future_needs_cd, by = c("cduid", "cdname")) %>% 
  mutate(rentable.square.feet.needed = courtrooms.needed * courtroom_size(courtrooms.needed)) %>% 
  ungroup()

saveRDS(courthouse_asset_with_projected_cd, "output/mag/courthouse_asset_with_projected_cd.rds")



# Calculating CD utilization level - 3 approaches tested
# Utilization at this stage refers to the ratio of needed courtrooms to current number of courtrooms
# Approach most realistic and accurate is summing business line courtroom needs by CD

## Approach 1 - Calculate an average utilization by CD and model year
cd_utilization_appr1 <- courthouse_asset_with_projected_cd %>% 
  mutate(utilization_appr1 = courtrooms.needed/number.of.courtrooms.2018)

## Approach 2 - Calculating utilization by weighting BL courtroom needs by BL courtroom hrs -- comment: this approach underestimated the need for courtrooms by order of magnitudes. It circumvents the results of the operational utilization graphs. not recommended for use
# cd_utilization_appr2 <- future_needs_cd_casetypes %>% 
#   group_by(cduid) %>% 
#   mutate(total = sum(courthours.needed), weight= courthours.needed/total, adjst_courtrooms_needed = courtrooms.needed*weight) %>% 
#   select(-total, -weight) %>% 
#   summarise(cdname = first(cdname), adj_cd_courtroom_needs = sum(adjst_courtrooms_needed)) %>% 
#   mutate(adjst_courtroom_needs_rounded = round(adj_cd_courtroom_needs,0)) %>% 
#   left_join(select(courthouse_asset_with_projected_cd, number.of.courtrooms.2018, cduid), by = "cduid") %>% 
#   mutate(utilization_appr2 =adjst_courtroom_needs_rounded/number.of.courtrooms.2018)

## Approach 3 - create a new 'general' operational utilization chart and assume each CD is one large courthouse:: comments: generates resonable estimates and close to the by casetype approach, however very slightly underestimates the results.


###create a general rate using weighted means of the total sums of BL CDs
# cd_utilization_appr3_utilization_chart <- future_needs_cd_casetypes %>% 
#   group_by(casetypes) %>% 
#   summarise(courthours.needed = sum(courthours.needed)) %>% 
#   left_join(utilization_rate, by = "casetypes") %>% 
#   summarise(base.hrs.dy.weighted = weighted.mean(base.hours.dy, courthours.needed), inflect.hrs.dy.weighted = weighted.mean(inflect.hours.dy, courthours.needed), max.hrs.dy.weighted = weighted.mean(max.hours.dy, courthours.needed), inflect.hours.weighted = weighted.mean(inflect.hours, courthours.needed), max.hours.weighted = weighted.mean(max.hours, courthours.needed)) %>% 
#   rename(base.hours.dy = base.hrs.dy.weighted, inflect.hours.dy = inflect.hrs.dy.weighted, max.hours.dy = max.hrs.dy.weighted, inflect.hours = inflect.hours.weighted, max.hours = max.hours.weighted)


###Apply the general rate to the sum of CD courthr totals
# cd_utilization_appr3 <- readRDS("cache/mag/future_cases_bl.rds") %>%
#   filter(year == MODEL_YEAR) %>% 
#   group_by(cduid, cdname, year) %>% 
#   summarise(court.hours = sum(court.hours)) %>% 
#   crossing(cd_utilization_appr3_utilization_chart) %>% 
#   mutate(op.utilization = 0) %>% 
#   mutate(op.utilization = ifelse(court.hours <= inflect.hours,
#                               base.hours.dy + (inflect.hours.dy - base.hours.dy) / (inflect.hours - 0 ) * court.hours,
#                               op.utilization)) %>% 
#   mutate(op.utilization = ifelse(court.hours > inflect.hours & court.hours <= max.hours,
#                                  inflect.hours.dy + ((max.hours.dy - inflect.hours.dy) * (court.hours - inflect.hours)                                                       / (max.hours - inflect.hours)),
#                                  op.utilization)) %>% 
#   mutate(op.utilization = ifelse(court.hours > max.hours,
#                                  max.hours.dy,
#                                  op.utilization)) %>% 
#   mutate(courtrooms.needed = round(court.hours / (op.utilization * OP_DAYS), digits = 2)) %>% 
#   left_join(select(courthouse_asset_with_projected_cd, number.of.courtrooms.2018, cduid), by = "cduid") %>% 
#   mutate(utilization_appr3 =courtrooms.needed/number.of.courtrooms.2018) %>% 
#   left_join(select(cd_utilization_appr1, utilization_appr1, cduid), by="cduid")

  
```

# Demand Distribution

- NOTE: this may not be necessary anymore after MAG Workshop
- All cases are assigned to a single courthouse because the buisness lines do NOT compete with each other
- Upon confirmation, the distribution of demand to competing courthouses can be removed.

```{r demand distribution}

# Using the 20/80 & 10/90 random/closest courthouse approach

CLOSEST_PERCENTAGE = 0.8
RANDOM_PERCENTAGE = 0.2
MODEL_YEAR = 2025
MODERNIZATION_FACTOR = tibble(casetypes = c('OCJcrim','SCJcrim','OCJfam','SCJfam','civil','smac'),
                              mod.factor = c(1, 1, 1, 1, 1, 1))
APPEARANCE_FACTOR = tibble(casetypes = c('OCJcrim','SCJcrim','OCJfam','SCJfam','civil','smac'),
                           app.factor = c(2, 2, 2, 2, 2, 2))
FIXED_COST = 10000000
VARIABLE_COST = 2000000

# Prepare the origin demand vector for selected year
treso_demand_origin <- readRDS("cache/mag/future_cases_bl_treso.rds") %>%
  rename(treso.id.por = treso_zone) %>% 
  filter(year == MODEL_YEAR) %>%
  select(-year, -population, -pop.rate) %>%
  left_join(MODERNIZATION_FACTOR, by = "casetypes") %>%
  mutate(court.hours = court.hours * mod.factor)

# New Courthouse
new_courthouse <- tibble(
  bid = "B99999",
  courthouse.lat = 43.90032, courthouse.long = 	-78.85863,
  name = "New Hospital",
  building.type = "Base Court",
  casetypes.serviced = c("civil"),
  courtrooms = 4,
  rentable.square.feet = 1000)

# Overlay with TRESO Shapefile to get the TRESO/CD Information
new_courthouse_xy <- create_court_xy(new_courthouse)
new_courthouse_overlay <- create_overlay(new_courthouse_xy, treso_shp, type = "court") %>% 
  left_join(treso_zone_system, by = c("treso.id.pos" = "treso_id")) %>% 
  left_join(new_courthouse, by = "bid") %>% 
  mutate_if(is.factor, as.character)

# Find out the treso zone and CD region
courthouse_asset_user <- bind_rows(courthouse_asset, new_courthouse_overlay)

# Create a list of courthouses available in each CD
courthouse_available <- courthouse_asset_user %>%
  unnest(casetypes = strsplit(casetypes.serviced, ",")) %>% 
  group_by(cduid, cdname, casetypes) %>% 
  summarise(treso.id.pos = paste(treso.id.pos, collapse = ","),
            bid = paste(bid, collapse = ","),
            name = paste(name, collapse = ","),
            building.type = paste(building.type, collapse = ","),
            courtrooms = paste(courtrooms, collapse = ",")
            )

# Join the list of courthouses available by CD to the origin demand
# NOTE: Sudbury (CD 3552) needs to be matched with Greater Sudbury (CD 3553) courts
# Distribute a percentage of TRESO demand to closest courthouse within the CD
# Distribute another percentage of TRESO demand weight by available courhouses within the CD
courthours_od <- treso_demand_origin %>%
  filter(court.hours != 0) %>% 
  mutate(cduid = ifelse(cduid == 3552, 3553, cduid),
         cdname = ifelse(cdname == "Sudbury", "Greater Sudbury / Grand Sudbury", cdname)) %>% 
  left_join(courthouse_available, by = c("cduid", "cdname", "casetypes")) %>% 
  unnest(treso.id.pos = strsplit(treso.id.pos, ","),
         bid = strsplit(bid, ","),
         name = strsplit(name, ","),
         building.type = strsplit(building.type, ","),
         courtrooms = strsplit(courtrooms, ",")) %>% 
  mutate(treso.id.por = as.numeric(treso.id.por),
         treso.id.pos = as.numeric(treso.id.pos),
         courtrooms = as.numeric(courtrooms)) %>% 
  # Merge with TRESO travel times
  left_join(travel_time_skim, by = c("treso.id.por", "treso.id.pos")) %>%
  # Identify the closest travel time pair by calculating a travel utility (e^(1/tt))
  mutate(travel.utility = exp(1/value)) %>%
  arrange(treso.id.por, casetypes, desc(travel.utility)) %>% 
  group_by(treso.id.por, casetypes) %>% 
  # Give a utility ranking for each combination of POR and casetypes
  mutate(travel.utility.ranking = row_number(),
         courtroom.weight = courtrooms / sum(courtrooms)) %>% 
  ungroup() %>% 
  # Flag the courthouses in the same TRESO zone servicing the same casetype that has the closest travel utility
  group_by(treso.id.por, casetypes, travel.utility) %>% 
  mutate(identical = ifelse(nchar(paste(unique(bid), collapse = ",")) > 6 && travel.utility.ranking == 1,
                            ceiling(nchar(paste(unique(bid), collapse = "")) / 6) , 0)) %>% 
  ungroup() %>% 
  # If the courthouse is the closest but also have a identical flag - change the travel utility ranking to 2
  mutate(travel.utility.ranking = ifelse(identical > 0, 2, travel.utility.ranking)) %>% 
  mutate(court.hours.to.closest = ifelse(travel.utility.ranking == 1, court.hours * CLOSEST_PERCENTAGE, 0),
         court.hours.to.random = court.hours * RANDOM_PERCENTAGE * courtroom.weight,
         court.hours.to.identical = ifelse(identical > 0, court.hours * CLOSEST_PERCENTAGE / identical, 0))

saveRDS(courthours_od, "output/mag/courthours_od.rds")

# Calculate the trip list by converting the hours to the courthouse with an user-defined
# appearance factor
SELECTED_COURTHOUSE_BID = "B20705"
appearance_trip_list <- courthours_od %>% 
  # filter(bid == SELECTED_COURTHOUSE_BID) %>% 
  group_by(treso.id.por, treso.id.pos, bid, casetypes) %>% 
  summarise(court.hours = sum(court.hours.to.closest, court.hours.to.random, court.hours.to.identical)) %>% 
  left_join(APPEARANCE_FACTOR, by = "casetypes") %>% 
  mutate(appearances = court.hours / app.factor) %>% 
  group_by(treso.id.por, treso.id.pos, bid) %>% 
  summarise(appearances = sum(appearances)) %>% 
  ungroup()

# Summarize the distributed demand to each courthouses and calculate the utilization for each courthouse
courthouse_asset_updated <- courthours_od %>% 
  group_by(bid, name, building.type, courtrooms, cduid, cdname, casetypes) %>% 
  summarise(court.hours.to.closest = sum(court.hours.to.closest),
            court.hours.to.random = sum(court.hours.to.random),
            court.hours.to.identical = sum(court.hours.to.identical)
            ) %>% 
  mutate(court.hours.distributed = court.hours.to.closest + court.hours.to.random + court.hours.to.identical) %>% 
  left_join(utilization_rate, by = "casetypes") %>% 
  mutate(op.utilization = 0) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed <= inflect.hours,
                              base.hours.dy + (inflect.hours.dy - base.hours.dy) / (inflect.hours - 0 ) * court.hours.distributed,
                              op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed > inflect.hours & court.hours.distributed <= max.hours,
                                 inflect.hours.dy + ((max.hours.dy - inflect.hours.dy) * (court.hours.distributed - inflect.hours)                                                       / (max.hours - inflect.hours)),
                                 op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed > max.hours,
                                 max.hours.dy,
                                 op.utilization)) %>% 
  mutate(courtrooms.needed = round(court.hours.distributed / (op.utilization * OP_DAYS), digits = 2)) %>% 
  mutate(actual.utilization = court.hours.distributed / courtrooms / OP_DAYS,
         utilization = actual.utilization / op.utilization) %>% 
  group_by(bid, name, building.type, cduid, cdname) %>% 
  summarise(courtrooms = first(courtrooms),
            courtrooms.needed = sum(courtrooms.needed),
            court.hours.distributed = sum(court.hours.distributed),
            utilization = sum(utilization)
            ) %>% 
  ungroup() %>% 
  mutate(cost = FIXED_COST + VARIABLE_COST * courtrooms.needed,
         standard.area = courtroom_size(courtrooms.needed)) %>% 
  select(bid, courtrooms, courtrooms.needed, court.hours.distributed, utilization, standard.area, cost)
```

# Create 2018 Observed Data

```{r observed 2018 court hours to all available courts}
cd_court_hours <- court_hours %>% 
  filter(year == 2018) %>% 
  group_by(cduid, cdname, casetypes) %>%
  summarise(court.hours = sum(value)) %>% 
  ungroup()

courthouse_2018_observed <- courthouse_asset %>% 
  unnest(casetypes = strsplit(casetypes.serviced, ",")) %>% 
  left_join(cd_court_hours, by = c("cduid", "cdname", "casetypes")) %>% 
  group_by(cduid, cdname, casetypes) %>% 
  mutate(courtroom.weight = courtrooms / sum(courtrooms)) %>% 
  ungroup() %>% 
  mutate(court.hours.distributed = court.hours * courtroom.weight) %>% 
  left_join(utilization_rate, by = "casetypes") %>% 
  mutate(op.utilization = 0) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed <= inflect.hours,
                              base.hours.dy + (inflect.hours.dy - base.hours.dy) / (inflect.hours - 0 ) * court.hours.distributed,
                              op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed > inflect.hours & court.hours.distributed <= max.hours,
                                 inflect.hours.dy + ((max.hours.dy - inflect.hours.dy) * (court.hours.distributed - inflect.hours)                                                       / (max.hours - inflect.hours)),
                                 op.utilization)) %>% 
  mutate(op.utilization = ifelse(court.hours.distributed > max.hours,
                                 max.hours.dy,
                                 op.utilization)) %>% 
  mutate(courtrooms.needed = round(court.hours.distributed / (op.utilization * OP_DAYS), digits = 2)) %>% 
  mutate(actual.utilization = court.hours.distributed / courtrooms / OP_DAYS,
         utilization = actual.utilization / op.utilization) %>% 
  group_by(bid, name, address, lat, long, courthouse.lat, courthouse.long, building.type, cduid, cdname) %>% 
  summarise(courtrooms = first(courtrooms),
            rentable.square.feet = first(rentable.square.feet),
            casetypes.serviced = first(casetypes.serviced),
            courtrooms.needed = sum(courtrooms.needed),
            court.hours.distributed = sum(court.hours.distributed),
            utilization = sum(utilization)
            ) %>% 
  ungroup()

saveRDS(courthouse_2018_observed, "output/mag/courthouse_2018_observed.rds")
```

## New Courthouse Cost Function
```{r court cost function}
# Read in cost function parameters from model user: cost per courtroom plus a startup cost applied to all courthouses regardless of size
user_input_cost_per_courtroom <- 30000000 # TODO: Replace this hard-coded value with user-input value from visualization
user_input_flat_courthouse_cost <- 20000000 # TODO: Replace this hard-coded value with user-input value from visualization

# Read in minimum and maximum size of courthouse (in number of beds) for potential new hospitals
#min_courthouse_size <- 1 # TODO: Replace this hard-coded value with user-input value from visualization
#max_courthouse_size <- 50 # TODO: Replace this hard-coded value with user-input value from visualization

# Read in list of new hospitals
# TODO: Replace this hard-coded tibble with user-input value from visualization
new_courthouse_list <- tibble(user_input_OCJcrim_courtrooms = c(3, 0), 
                        user_input_SCJcrim_courtrooms = c(2, 0), 
                        user_input_OCJfam_courtrooms = c(0, 0), 
                        user_input_SCJfam_courtrooms = c(0, 0), 
                        user_input_civil_courtrooms = c(0, 4), 
                        user_input_smac_courtrooms = c(0, 3), 
                        user_input_courtrooms_total = c(5, 7)) 

# Calculate costs
new_courthouse_list_complete <- new_courthouse_list %>% 
  mutate(cost_2018 = round(user_input_courtrooms_total * user_input_cost_per_courtroom + user_input_flat_courthouse_cost, 0))

```

## Cost Annualization 
- Inflate cost for facility construction based on construction year; calculate total cost over time period of interest
```{r}
user_input_inflation <- 0.02 # TODO: Replace this hard-coded value with user-input value from visualization
user_input_scenario_year <- 2025 # TODO: Replace this hard-coded value with user-input value from visualization

currency_year = 2018 # Units of costs; set to 2018 since cost function from EDU assumed to be set in 2018$
current_year = 2019 # Current year used to calculate construction timespan; should be parameterized and put up front for user to choose

annualized_inflated_cost <- construction_cost(user_input_inflation, user_input_scenario_year, currency_year,
                                              current_year, new_courthouse_list_complete)
```

# Linear Regression

- Commented out due to being non-essential to current modelling approach, but can be used for further exploration
- Conducted to test whether there are any linear relationships between age groups and certain court hours

```{r explore linear regression}
# library(perturb)

# treso_2016_population <- readRDS("input/mag/treso_2016_population_mag.rds") %>% 
#   left_join(treso_zone_system, by = c("treso_zone" = "treso_id")) %>% 
#   select(treso_zone, age.group, population, cduid, cdname) %>% 
#   mutate(cdname = as.character(cdname)) %>% 
#   group_by(cduid, cdname, age.group) %>% 
#   summarise(population = sum(population)) %>% 
#   ungroup() %>% 
#   spread(key = "age.group", value = "population") %>% 
#   mutate(adult = adult.20s + adult.30s + adult.40s + adult.50s + adult.60s) %>% 
#   mutate(senior = senior.70s + senior.80s) %>% 
#   mutate(majority = adult + senior) %>% 
#   mutate(minor = child + youth + young.adult) %>% 
#   mutate(total = majority + minor)
# 
# court_hours_2016 <- court_hours %>% 
#   filter(year == 2016) %>% 
#   select(-lat, -long, -csduid, -csdname, -area, -region, -courthouse.lat, -courthouse.long) %>% 
#   group_by(cduid, cdname, casetypes) %>%
#   summarise(value = sum(value)) %>% 
#   ungroup() %>% 
#   spread(key = "casetypes", value = "value") %>% 
#   left_join(treso_2016_population, by = c("cduid", "cdname")) %>% 
#   select(-cduid, -cdname) %>% 
#   mutate_if(is.numeric, list(ln = log1p, sqrt = sqrt, log = log10, sq = ~.^2, scale = scale)) 

```

```{r building models - OCJcrim}
# r1 <- resid(lm(young.adult_ln ~ adult.20s_ln, data = court_hours_2016))
# mod <- lm(OCJcrim_sqrt ~ adult.20s_ln + r1, data = court_hours_2016)
# summary(mod)
# 
# plot(mod)
```

```{r building models - SCJcrim}
# r1 <- resid(lm(adult.30s_sqrt ~ adult.20s_sqrt, data = court_hours_2016))
# mod <- lm(SCJcrim_sqrt ~ adult.20s_sqrt + r1, data = court_hours_2016)
# summary(mod)
# plot(mod)
```

```{r building models - OCJfam}
# mod <- lm(OCJfam ~ adult.30s_sqrt + adult.50s_sqrt + child, data = court_hours_2016)
# summary(mod)
# plot(mod)
# 
# # Comment - it was found that to have a high R^2 and proper signs with the `child` term, the only transformation is the non-transformed version.
```

```{r building models - SCJfam}
# mod <- lm(SCJfam ~ youth_sqrt + senior.70s, data = court_hours_2016)
# summary(mod)
# plot(mod)
```

```{r building models - civil}

# mod <- lm(civil_sqrt ~ young.adult_sqrt + adult.20s_sqrt + senior.80s_sqrt, data = court_hours_2016)
# summary(mod)
# print(colldiag(mod))
# plot(mod)
```

```{r building models - smac}
# rMod <- residuals(lm(adult.30s ~ adult.20s, data = court_hours_2016))
# mod <- lm(smac ~ adult.20s_sqrt + rMod, data = court_hours_2016)
# cd <- colldiag(mod)
# cd
# 
# mod <- lm(smac ~ adult.20s_sqrt + adult.30s_sqrt, data = court_hours_2016)
# cd <- colldiag(mod)
# cd
# 
# #plot(mod)

```

# Decision Making
```{r decision making prep}


U_AVG <- 0.8 # User input

# Calculate the anticipated increase in court hour demand to yield the U_AVG output. For example, 
#U=D/C => C = D/U => C = courtroom hour demand / U_AVG. This intermediate step does not alter demand 
#values, only uses them to attain the U_AVG.
future_cases_bl_cd_util <- readRDS("cache/mag/future_cases_bl_treso.rds") %>% 
  left_join(select(treso_zone_system, csduid, treso_id), by = c("treso_zone" = "treso_id")) %>% 
  group_by(cduid, year, casetypes) %>%
  summarize(cd.court.hours = sum(court.hours)) %>% 
  mutate(util.court.hours = cd.court.hours * (1/U_AVG))

# Read the casetype utilization rate table
utilization_rate <- readRDS("cache/mag/utilization_rate.rds")

# Read the existing 2018 asset list for number of courtrooms existing
courthouse_asset_cd_level <- courthouse_asset %>%
  group_by(cduid, cdname) %>% 
  summarise(number.of.courtrooms.2018 = sum(courtrooms)) %>% 
  ungroup()

# Calculate future court hours needed based on the business line specific utilization increase characteristics
future_needs_util <- future_cases_bl_cd_util %>%
  filter(year == MODEL_YEAR) %>% 
  left_join(utilization_rate, by = "casetypes") %>% 
  mutate(op.utilization = 0) %>% 
  mutate(op.utilization = ifelse(util.court.hours <= inflect.hours,
                              base.hours.dy + (inflect.hours.dy - base.hours.dy) / (inflect.hours - 0 ) * util.court.hours,
                              op.utilization)) %>% 
  mutate(op.utilization = ifelse(util.court.hours > inflect.hours & util.court.hours <= max.hours,
                                 inflect.hours.dy + ((max.hours.dy - inflect.hours.dy) * (util.court.hours - inflect.hours)                                                       / (max.hours - inflect.hours)),
                                 op.utilization)) %>% 
  mutate(op.utilization = ifelse(util.court.hours > max.hours,
                                 max.hours.dy,
                                 op.utilization)) %>% 
  mutate(util.courtrooms.needed = round(util.court.hours / (op.utilization * OP_DAYS), digits = 2)) %>% 
  select(-base.hours.dy, - inflect.hours.dy, -max.hours.dy, -inflect.hours, -max.hours, -op.utilization, -cd.court.hours) 

saveRDS(future_needs_util, "cache/mag/future_needs_util.rds")

# Calculate the difference between number of courtrooms needed and number of courtrooms existing.
courtroom_net <- future_needs_util %>% 
  group_by(cduid) %>% 
  summarise(util.courtrooms.needed = sum(util.courtrooms.needed)) %>% 
  left_join(select(courthouse_asset_cd_level, cduid, number.of.courtrooms.2018), by = "cduid") %>% 
  mutate(number.of.courtrooms.2018 - util.courtrooms.needed)


```


