---
title: "edu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(rgeos)
library(rgdal)
library(sp)
library(foreach)
library(doParallel)
library(tictoc)
source("Data modelling functions.2019-03-28.R")
```


## Import relavent files
```{r Import files}
# Treso shapefiles
treso_shp <- readOGR(dsn = "input",
                     layer = "TRESO_Zones_SocioData_Gatineau_LCC")

# 2011 hhld and population data by treso zone
hhld_df <- data.table::fread("input/households_2011.csv.gz")
pop_df <- data.table::fread("input/persons_2011.csv.gz")

# Definition files
school_board_def <- read_csv("input/school_board_types.csv")
treso_zone_def <- read_csv("input/treso_zone_system.csv")

# Students within 90th percentile
student_travel <- read_csv("output/student_travel.csv")

# School's historical ADE
school_ade <- read_csv("output/school_historical_ade.csv")


```

## Data cleaning

- Get the coordinates for schools and students
- Get the projection from the TRESO shapefile
- Transform the coordinates to use the same projection system as TRESO shapefile
```{r data cleaning}
# Convert the student dataframe into SpatialPointsDataframe
student_spdf <- select(student_travel, c(X1, studentLat, studentLong, dist, schoolName.x, Enrolment)) %>%
  rename(
    id = X1,
    lat = studentLat,
    long = studentLong,
    euclidean_dist = dist,
    school_name = schoolName.x,
    enrolment = Enrolment
    )

coordinates(student_spdf) <- c('long', 'lat')

# Convert the school dataframe into SpatialPointsDataframe
school_spdf <- select(student_travel, c(schoolName.x, schoolLat.x, schoolLong.x, DSBINDEX.x, SFIS, percDist)) %>%
  group_by(schoolName.x) %>%
  summarise(schoolLat = first(schoolLat.x), schoolLong = first(schoolLong.x), 
            dsb = first(DSBINDEX.x), percDist = first(percDist), sfis = first(SFIS)) %>%
  transform(id = 1:nrow(.)) %>%
  rename(
    lat = schoolLat,
    long = schoolLong,
    school_name = schoolName.x,
    catchment_dist = percDist
  )
coordinates(school_spdf) <- c('long', 'lat')

# Project the student and school points from lat/long to TRESO's LCC specification
proj4string(student_spdf) <- CRS("+proj=longlat +datum=WGS84")
proj4string(school_spdf) <- CRS("+proj=longlat +datum=WGS84")
treso_projarg = treso_shp@proj4string@projargs

student_xy <- spTransform(student_spdf, CRS(treso_projarg))
school_xy <- spTransform(school_spdf, CRS(treso_projarg))
```

- Combine the population, household information
- Summarize by TRESO zones
```{r combine socio-economic data}

# Clean hhld_df
hhld_tb <- select(hhld_df, hhid, treso_zone, income) %>%
  as.tibble()

# Clean pop_df and join with hhld
socio_economic_tb <- select(pop_df, hhid, treso_zone, hdgree, nocs_11, work_status, school_status) %>%
  as.tibble() %>%
  left_join(hhld_tb, by = c("hhid" , "treso_zone"))

# Summarize into TRESO Zones
treso_tb <- socio_economic_tb %>%
  group_by(treso_zone) %>%
  summarize(
    n_pop = n(),
    n_hhlds = length(unique(hhid)),
    n_ft = length(treso_zone[work_status == "full_time"]),
    n_pt = length(treso_zone[work_status == "part_time"]),
    n_unemp = length(treso_zone[work_status == "unemployed"]),
    mean_income = mean(income, na.rm = TRUE),
    occu_management = length(treso_zone[nocs_11 == 1]),
    occu_business = length(treso_zone[nocs_11 == 2]),
    occu_science = length(treso_zone[nocs_11 == 3]),
    occu_health = length(treso_zone[nocs_11 == 4]),
    occu_public = length(treso_zone[nocs_11 == 5]),
    occu_recreation = length(treso_zone[nocs_11 == 6]),
    occu_sales = length(treso_zone[nocs_11 == 7]),
    occu_trades = length(treso_zone[nocs_11 == 8]),
    occu_production = length(treso_zone[nocs_11 == 9]),
    occu_manufacturing = length(treso_zone[nocs_11 == 10]),
    occu_notapplicable = length(treso_zone[nocs_11 == 99]),
    attend_school = length(treso_zone[school_status == 2]),
    deg_none = length(treso_zone[hdgree == 1]),
    deg_hs = length(treso_zone[hdgree == 2]),
    deg_trades = length(treso_zone[hdgree == 3]),
    deg_ra = length(treso_zone[hdgree == 4]),
    deg_col = length(treso_zone[hdgree == 5]),
    deg_uni = length(treso_zone[hdgree == 6]),
    deg_ugrad = length(treso_zone[hdgree == 7]),
    deg_grad = length(treso_zone[hdgree == 8]),
    deg_na = length(treso_zone[hdgree == 88 | hdgree == 99])
  )

# Clean global variables
rm(hhld_tb, socio_economic_tb)
```

### Plot TRESO zones, with school and student 

```{r plotting shapefiles}

# Convert the SpatialDataframe to normal Dataframe for ggplot()
treso_shp_df <- fortify(treso_shp) %>%
  rename(
    x = long,
    y = lat
  )

school_xy_df <- data.frame(school_xy) %>%
  rename(
    x = long,
    y = lat
  )

student_xy_df <- data.frame(student_xy) %>%
  rename(
    x = long,
    y = lat
  )


treso_plot <- ggplot() +
  geom_polygon(data = treso_shp_df, mapping = aes(x = x, y = y, group = group), 
               fill = 'grey40', color = 'grey90', alpha = 1, size = 0.1) +
  geom_point(data = student_xy_df, mapping = aes(x = x, y = y), color = "red", alpha = 0.1, size = 0.01) +
  geom_point(data = school_xy_df, mapping = aes(x = x, y = y), color = "blue", alpha = 0.1, size = 0.01, shape = 2) +
  labs(x = "", y = "", title = "Students and schools in Ontario") +
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        plot.title = element_text(lineheight = 0.8, face = "bold", vjust = 1)) +
  coord_equal(ratio = 1)
ggsave(treso_plot, file = "output/treso_with_points.png", width = 10, height = 10)

# Clean up the global environments
rm(school_xy_df, treso_shp_df, student_xy_df)
```

## Spatial gymnastics
- Using over() to determine the treso zone number for each student/school point
- Using treso travel time skims (*need to obtain*), plot the observed TLFD of students and their destination.

```{r spatial gymnastics}

# Find the treso zones which the student points layover
student_overlay <- over(student_xy, treso_shp, returnList = FALSE) %>%
  cbind(euclidean_dist = student_xy@data$euclidean_dist,
        student_id = student_xy@data$id,
        school_name = student_xy@data$school_name) %>%
  as.tibble() %>%
  select(Treso_ID, euclidean_dist, student_id, school_name) %>%
  rename(
    treso_id_por = Treso_ID
  )

# Find the treso zones which the school points layover
school_overlay <- over(school_xy, treso_shp, returnList = FALSE) %>%
  cbind(catchment_dist = school_xy@data$catchment_dist,
        dsb = school_xy@data$dsb,
        school_name = school_xy@data$school_name,
        sfis = school_xy@data$sfis) %>%
  as.tibble() %>%
  select(Treso_ID, catchment_dist, dsb, school_name) %>%
  rename(
    treso_id_pos = Treso_ID
  )


# Join the POR and POS together for distance matrix
observed_por_pos <- left_join(student_overlay, school_overlay, by = "school_name")
```

- Using gbuffer() to determine the treso zones that the 90th percentile distance touches for each school
- Setting up the multi-core to do `foreach()` loop is worth it once you are doing more than 100 fields
  - 100 fields: 29s compared to 19s
  - 1000 fields: 252s compared to 116s

``` {r buffer treso zones to schools}
buffered_points <- gBuffer(school_xy, width = school_xy@data$catchment_dist * 1000, byid = TRUE)

tic("Multi-Core Buffering")
# Start multi-core clustering for performance
cores = parallel::detectCores()
cl <- parallel::makeCluster(cores[1] - 1)
registerDoParallel(cl)

# foreach() loop to find the TRESO zones touching the school buffers
datalist <- list()
datalist <- foreach(i = 1:nrow(buffered_points@data), .packages=c("sp", "tidyverse", "rgeos")) %dopar% {
  
  # Select the buffer
  buffer <- buffered_points[buffered_points@data$id == i, ]

  # Get all TRESO zones that touches the buffer with over(), gIntersect() could be another solution
  treso_overlay <- over(treso_shp, buffer, returnList = FALSE) %>%
    cbind(treso_id = treso_shp@data$Treso_ID,
          households = treso_shp@data$TotHH,
          persons = treso_shp@data$TotPers) %>%
    as.tibble() %>%
    select(school_name, sfis, dsb, catchment_dist, treso_id, households, persons) %>%
    drop_na()

  # Save each tibble in a list of tibbles
  datalist[[i]] <- treso_overlay
}

stopCluster(cl)
toc()

# Write the datalist to csv
datalist_df <- rbindlist(datalist)
write.csv(datalist_df, "output/school_catchment_treso_zones.csv", row.names = FALSE)
```

- Test that `gBuffer()` is returning intended results.

```{r test buffer}
# Select a specific buffer
school_of_interest <- "Marc Garneau CI"
buffer <- buffered_points[buffered_points@data$school_name == school_of_interest, ]

# Get all TRESO zones that touches the buffer with over(), gIntersect() could be another solution
treso_overlay <- over(treso_shp, buffer, returnList = FALSE) %>%
  cbind(treso_id = treso_shp@data$Treso_ID,
        households = treso_shp@data$TotHH,
        persons = treso_shp@data$TotPers) %>%
  as.tibble() %>%
  select(school_name, sfis, dsb, catchment_dist, treso_id, households, persons) %>%
  drop_na()

# Select the TRESO zones of interest
treso_zones_interest <- treso_shp[is.element(treso_shp@data$Treso_ID, treso_overlay$treso_id), ] %>%
  fortify() %>%
  rename(
    x = long, 
    y = lat
  )

# Convert buffer into df
buffer_df <- fortify(buffer) %>%
  rename(
    x = long, 
    y = lat
  )

school_xy_df <- data.frame(school_xy) %>%
  rename(
    x = long,
    y = lat
  )

# Plot the school, buffer and the selected TRESO zones
ggplot() +
  geom_polygon(data = treso_zones_interest, mapping = aes(x = x, y = y, group = group), 
               fill = 'grey40', color = 'grey90', alpha = 1, size = 0.1) +
  geom_polygon(data = buffer_df, mapping = aes(x = x, y = y, group = group),
               fill = 'red', color = 'red', alpha = 0.5) +
  geom_point(data = filter(school_xy_df, school_name == school_of_interest), mapping = aes(x = x, y = y), 
               fill = "blue", alpha = 0.5, size = 10) +
  labs(x = "", y = "", title = "Test") +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold", vjust = 1)) +
  coord_equal(ratio = 1)

# Clean up global environment variables
rm(school_of_interest, buffer, treso_overlay, treso_zones_interest, buffer_df, school_xy_df)
```

## Linear Regression

- Combine the the datalist of buffered TRESO zones with the tibble containing socio-economic data
- Combine with the historical ADE and utilization data by school

```{r prepare data for linear regression}
# Combine the school's buffered TRESO zones with socio-economic info and group_by school name
# First summarise data that is calculated with mean
# I think using purr, one could summarise different columns with different functions
school_tb_temp <- as.tibble(datalist_df) %>%
  left_join(select(treso_tb, treso_zone, mean_income), by = c("treso_id" = "treso_zone")) %>%
  left_join(select(school_board_def, dsb, board_type_name), by = c("dsb")) %>%
  replace(is.na(.), 0) %>%
  group_by(school_name) %>%
  summarise(
    mean_income = mean(mean_income),
    dsb = first(dsb),
    sfis = first(sfis),
    catchment_dist = first(catchment_dist),
    board_type_name = first(board_type_name)
  )

# Then, combine with data that is calculated with sum
school_tb <- as.tibble(datalist_df) %>%
  left_join(treso_tb, by = c("treso_id" = "treso_zone")) %>%
  replace(is.na(.), 0) %>%
  group_by(school_name) %>%
  summarise_at(
    .vars = vars(starts_with("n_"), starts_with("occu_"), starts_with("deg_"), "attend_school"),
    .funs = c(sum = "sum")
  ) %>%
  left_join(school_tb_temp, by = "school_name")

# Summarize to obtain with the utility info by year
# Using reshape2, one could combine these two steps into one, but I wanted to keep the requirements down. 
school_util_tb <- school_ade %>%
  filter(Status == "Open") %>%
  select(Sch_yr, SFIS_ID, FacilityName, utilization) %>%
  rename(
    year = Sch_yr,
    sfis = SFIS_ID,
    school_name = FacilityName
  ) %>%
  spread(year, utilization) %>%
  rename_at(vars(starts_with("20")), ~ paste0("util_", .))

# Summarize to obtain with the ADE info by year
school_ade_tb <- school_ade %>%
  filter(Status == "Open") %>%
  select(Sch_yr, Panel, SFIS_ID, FacilityName, ADE) %>%
  rename(
    year = Sch_yr,
    panel = Panel,
    sfis = SFIS_ID,
    school_name = FacilityName,
    ade = ADE
  ) %>%
  spread(year, ade) %>%
  rename_at(vars(starts_with("20")), ~ paste0("ade_", .)) %>%
  left_join(school_util_tb, by = c("sfis", "school_name")) %>%
  replace(is.na(.), 0)

# Combine School's ADE and utilization info with School's buffered zones
school_tb <- school_tb %>%
  left_join(school_ade_tb, by = c("sfis", "school_name"))

# Clean up the global environments
rm(school_tb_temp, school_util_tb)
```

- Segment the schools by board (at least)
- Build linear regression 
- Explore other segmentations

```{r linear regression}

# Select schools with:
# - 2017 ADE value
# - Elementary
# - English Public
school_tb_filtered <- filter(school_tb, ade_2017 > 0, panel == "Elementary", board_type_name == "English Public")

# Build linear regression data base with various transformed x, y variables
linreg <- school_tb_filtered %>%
  select(ade_2017, starts_with("n_"), starts_with("occu_"), starts_with("deg_"), attend_school_sum, 
         mean_income) %>%
  mutate_all(
    funs(ln = log1p(.), sqrt = sqrt(.), sq = .^2)
  ) %>%
  cbind(select(school_tb_filtered, sfis, school_name))

# Exploring the data relationships
ggplot(linreg, aes(y = ade_2017_ln, x = occu_sales_sum_sqrt)) +
  geom_point(alpha = 0.5, color = "#F8766D") +
  geom_smooth(method = "lm", se = FALSE)

# Attemp a fit
fit_base <- lm(ade_2017_ln ~ occu_sales_sum_sqrt, data = linreg)
summary(fit_base)
plot(fit_base)
```


## Plot 90th percentile with straight line distances


- Convert this to use the TRESO distance once we obtain the skims.
- The code also use `uncount()` to convert the enrollment number to individual entries in the tibble

```{r tlfd plots by boards, fig.width = 10, fig.height = 7, echo = FALSE}
student_tlfd <- student_travel %>% 
  select(DSBINDEX.x, dist, Enrolment) %>%
  rename(dsb = DSBINDEX.x) %>%
  mutate(student_id = row_number()) %>%
  left_join(select(school_board_def, dsb, board_type_name), by = "dsb") %>%
  left_join(select(student_overlay, student_id, treso_id_por), by = "student_id") %>%
  rename(treso_id = treso_id_por) %>%
  left_join(select(treso_zone_def, treso_id, area), by = "treso_id") %>%
  # expand the rows based on enrollment number
  uncount(Enrolment)

ggplot(student_tlfd, aes(x = dist, color = as.factor(board_type_name), fill = as.factor(board_type_name))) +
  geom_freqpoly(stat = "bin", binwidth = 0.1, size = 1) + 
  facet_wrap(vars(board_type_name), scales = "free_y") +
  labs(title = "Euclidean Distance", x = "distance (km)", color = "Board Type") +
  theme_grey() +
  coord_cartesian(xlim = c(-1, 10))
```

```{r tlfd plots by regional areas, fig.width = 10, fig.height = 7, echo = FALSE}
ggplot(student_tlfd, aes(dist, color = as.factor(area), fill = as.factor(area))) +
  # geom_area(stat = "bin", binwidth = 0.1) +
  geom_freqpoly(stat = "bin", binwidth = 0.1, size = 1) + 
  facet_wrap(vars(board_type_name), scales = "free_y") +
  labs(title = "Euclidean Distance", x = "distance (km)", color = "Geographic Location") +
  theme_grey() +
  coord_cartesian(xlim = c(-1, 10))
```