---
title: "school_market_share"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(weights)
library(RODBC)
library(scales)
source("helpers.R")
```

# School Market Share Analysis - 2007 to 2011
## Education Data cleaning
 
- Clean school capacity and historical demand data

```{r education data cleaning, echo=FALSE}
# The school capacity dataset contains the correct lat/long for schools
schoolEnrol <- read.csv("input/edu/school_capacity.csv", stringsAsFactors = FALSE)
schoolEnrolSlim <- schoolEnrol %>%
  select(year = Sch_yr, 
         dsb.index = DSB_Index, 
         board.name = BoardName, 
         panel = Panel, 
         sfis = `SFIS_ID`,
         school.name = `FacilityName`, 
         ade = ADE,
         otg = OTG, 
         school.lat = `Latitude`, 
         school.long = `longitude`, 
         status = Status, 
         dataset.id = dataSetID) %>%
  arrange(year, dsb.index, panel, sfis) %>%
  mutate(utilization = ifelse(otg != 0, ade / otg, 0)) %>%
  filter(year == 2007 | year == 2011 | year == 2017, status == "Open", dataset.id == 2)
  # Remove schools with capacity of 0
  #filter(otg != 0, ade != 0) %>%

# Clean global variables
rm(schoolEnrol2011, schoolEnrol2017)
```

## Calculate 2007 Market Share by School
- Grouped by DSB Index, Panel

```{r}
schoolEnrol2007 <- schoolEnrolSlim %>%
  filter(year == 2007) %>%
  group_by(dsb.index, panel) %>%
  mutate(board.enrol = sum(ade)) %>%
  mutate(market.share2007 = ade / board.enrol) %>%
  ungroup()
```

## Join 2007 Market Shares to 2011 and 2017 Enrolment Data
- Estimates enrolment by school in 2011 and 2017 based on market share of that school within its board in 2007, using 2011 and 2017 board enrolment projections

```{r}
schoolEnrol2011 <- schoolEnrolSlim %>%
  filter(year == 2011) %>%
  left_join(select(schoolEnrol2007, sfis, market.share2007), by = c('sfis')) %>%
  group_by(year, dsb.index, panel) %>%
  mutate(board.enrol = sum(ade)) %>%
  ungroup() %>%
  mutate(enrol2011.by2007MS = market.share2007 * board.enrol) %>%
  mutate(MSerror2011 = abs(ade - enrol2011.by2007MS) / ade, MSerrorPercent2011 = 
         percent(abs(ade - enrol2011.by2007MS) / ade, accuracy = 1)) %>%
  
  # Calculate unweighted mean error
    # Find average enrolment differences for each year
    # Omit NaN and NA values using na.rm = TRUE
    # Omit Inf values by converting them to NA, and therefore caught by na.rm = TRUE
  group_by(panel) %>%
  mutate(avg2011EnrolDiff = mean(MSerror2011*is.finite(MSerror2011), na.rm = TRUE)) %>%
  ungroup()

# Repeat above for 2017
schoolEnrol2017 <- schoolEnrolSlim %>%
  filter(year == 2017) %>%
  left_join(select(schoolEnrol2007, sfis, market.share2007), by = c('sfis')) %>%
  group_by(year, dsb.index, panel) %>%
  mutate(board.enrol = sum(ade)) %>%
  ungroup() %>%
  mutate(enrol2017.by2007MS = market.share2007 * board.enrol) %>%
  mutate(MSerror2017 = abs(ade - enrol2017.by2007MS) / ade, MSerrorPercent2017 = 
         percent(abs(ade - enrol2017.by2007MS) / ade, accuracy = 1)) %>%
  
  # Calculate unweighted mean error
    # Find average enrolment differences for each year
    # Omit NaN and NA values using na.rm = TRUE
    # Omit Inf values by converting them to NA, and therefore caught by na.rm = TRUE
  group_by(panel) %>%
  mutate(avg2017EnrolDiff = mean(MSerror2017*is.finite(MSerror2017), na.rm = TRUE)) %>%
  ungroup()
  
```

## Join 2011 Results to 2017 Results and Compare; Save / output results

```{r}
schoolEnrol20112017 <- schoolEnrol2011 %>%
  select(dsb.index, sfis, school.name, panel, ade2011 = ade, enrol2011.by2007MS) %>%
  left_join(select(schoolEnrol2017, sfis, ade, enrol2017.by2007MS), by = c('sfis')) %>%
  rename(ade2017 = ade) %>%
  mutate(enrol2011diff = ade2011 - enrol2011.by2007MS, enrol2017diff = ade2017 - enrol2017.by2007MS) %>%
  arrange(sfis) %>%
  saveRDS("cache/edu/school_2007_market_share.rds")

```